### 2009-01-28 trunk/Papers/rmrs-08/rmrs.tex
in my view, beta

In this paper, we assume for simplicity that the base language is as
in \mrs; essentially, $\tau$ becomes the structure tree of a formula
of predicate logic.  We assume that $\Sigma$ is a ranked signature
consisting of the symbols of predicate logic: a unary constructor
$\neg$ and binary constructors $\wedge$, $\rightarrow$, etc.; a set of
3-place quantifier symbols such as $\sem{\_every\_q\_1}$ and
$\sem{\_some\_q\_1}$ (with the children being the bound variable, the
restrictor, and the scope); and constructors of various arities for
the predicate symbols; e.g., $\sem{\_chase\_v\_1}$ is of arity 3.  Other
base languages may require a different signature $\Sigma$ and/or a
different mapping between formulas and trees; the only strict
requirement we make is that the signature contains a binary
constructor $\wedge$ to represent conjunction.  We write $\Sigma_i$
for the set of all constructors in $\Sigma$ with arity $i$.  We
will follow the typographical convention that non-logical symbols in
$\Sigma$ are written in $\mathsf{sans{-}serif}$, as opposed to the
\rmrs\ predicate symbols like $\sempred{\_cat\_n}$ and
$\sempred{\_cat\_n\_1}$.

The models of {\sc rmrs} are then defined to be finite constructor
trees:
\begin{definition}\label{defn:models}
  A {\em finite constructor tree} $\tau$ is a function $\tau:D
  \rightarrow \Sigma$ such that $D$ is a \emph{tree domain} (i.e., a
  subset of $\N^*$ which is closed under prefix and left sibling) and
  the number of children of each node $u \in D$ is equal to the arity
  of $\tau(u)$.  
\end{definition}

@@

\begin{definition}\label{defn:dominance}
  $u \dom v$ iff $u$ is a prefix of $v$, i.e.\ the node $u$ is equal
  to or above
  the node $v$ in the tree.  (This is the dominance relation familiar
  from \newcite{egg:etal:2001}.)

  $u \wedgedom v$ iff $u \dom v$, and all symbols on the path from $u$
  to $v$ (not including $v$) are $\wedge$.
\end{definition}

@@

In the context of robust semantic processing, this is a desirable
feature of \rmrs, because it means that when we enrich an \rmrs\
obtained from a shallow processor with more semantic
information---such as the relation symbols introduced by
syntactic 
constructions such as appositives, noun-noun compounds and free
adjuncts---we don't change the set of models;
we only \emph{restrict} the set of models further and further towards
the semantic representation we are trying to reconstruct.
Furthermore, it has been shown in the literature that a
dominance-constraint style semantics for underspecified
representations gives us more room to manoeuvre when developing
efficient solvers than an \mrs-style semantics
\cite{Althaus_etal:JoA}.

@@

\begin{definition}  \label{defn:solved-form-of}
  The \emph{syntactic dominance relation} in an {\sc  rmrs} $\varphi$ is the
  reflexive, transitive closure of the binary relation $$
\begin{array}{r c l}
\{(X,Y) & \mid &  \mbox{$\varphi$ contains $X \dom Y$ or}\\
&& \mbox{$ARG_S(X,Y)$ for some
    $S$}\}
\end{array}
$$  
We write $D(\varphi)$ for the syntactic dominance
  relation of $\varphi$.

@@

Let's look at an example for all these definitions.  All the \rmrs s
presented in Section~\ref{sec:motivation} (replacing $\qeq$ by
$\dom$), are in solved form; this is least obvious for
(\ref{ex:cat-erg}), but becomes clear once we notice that no label is
on the right-hand side of two dominance atoms.  However, the model
constructed in the proof of
Prop.~\ref{prop:solved-forms-are-satisfiable} looks a bit like
Fig.~\ref{fig:fat-black-cat}; both models are problematic in several
ways and in particular contain an unbound variable $y$ even though
they also contains a quantifier that binds $y$.  If we restrict the
class of models to those in which such variables are bound (as
\newcite{copestake:etal:2005} do), we can enforce that the quantifiers
outscope their bound variables without changing models of the \rmrs\
further---i.e., we add the atoms $h_3 \dom l_5$ and $h_8 \dom l_5$.
Fig.~\ref{fig:fat-black-cat} is no longer a model for the extended
\rmrs, which in turn is no longer in solved form because the label
$l_5$ is on the right-hand side of two dominance atoms.  Instead, it
has the following two solved forms:

@@

So far we have based the syntax and semantics of \rmrs\ on the
dominance relation from \newcite{egg:etal:2001} rather than the qeq
relation from \newcite{copestake:etal:2005}.  This is partly because
dominance is a weaker relation than qeq: If a dependency parser links
a determiner to a noun and this noun to a verb, then we can use
dominance but not qeq to represent that the predicate introduced by
the verb is outscoped by the quantifier introduced by the determiner
(see earlier discussion).  However, it is very straightforward to add
qeq to the syntax of the language and to extend the model theory
accordingly.  This extension adds a new atom $X \dom Y$ to
Def.~\ref{defn:rmrs-syntax}, and $\tau,\alpha,g,\sigma$ will satisfy
$X \dom Y$ iff $\alpha(X)$ and $\alpha(Y)$ are separated by zero or more
nodes in $\tau$, each of which has a rightmost child that is a
quantifier. All the above propositions about solved forms still hold
if ``dominance'' is replaced with ``qeq''.

>>>>

In this paper, we assume for simplicity that the base language is as
in \mrs; essentially, $\tau$ becomes the structure tree of a formula
of predicate logic.  We assume that $\Sigma$ is a ranked signature
consisting of the symbols of predicate logic: a unary constructor
$\neg$ and binary constructors $\wedge$, $\rightarrow$, etc.; a set of
3-place quantifier symbols such as $\sem{\_every\_q\_1}$ and
$\sem{\_some\_q\_1}$ (with the children being the bound variable, the
restrictor, and the scope); and constructors of various arities for
the predicate symbols; e.g., $\sem{\_chase\_v\_1}$ is of arity 3.
Other base languages may require a different signature $\Sigma$ and/or
a different mapping between formulas and trees; the only strict
requirement we make is that the signature contains a binary
constructor $\wedge$ to represent conjunction.  We write $\Sigma_i$
and $\Sigma_{\geq i}$ for the set of all constructors in $\Sigma$ with
arity $i$ and at least $i$, respectively.  We will follow the
typographical convention that non-logical symbols in $\Sigma$ are
written in \textsf{sans-serif}, as opposed to the \rmrs\ predicate
symbols like $\sempred{\_cat\_n}$ and $\sempred{\_cat\_n\_1}$.

The models of {\sc rmrs} are then defined to be finite constructor
trees (see also \cite{egg:etal:2001}):
\begin{definition}\label{defn:models}
  A {\em finite constructor tree} $\tau$ is a function $\tau:D
  \rightarrow \Sigma$ such that $D$ is a \emph{tree domain} (i.e., a
  subset of $\N^*$ which is closed under prefix and left sibling) and
  the number of children of each node $u \in D$ is equal to the arity
  of $\tau(u)$.  
\end{definition}

@@

\begin{definition}\label{defn:dominance}
  $u \dom v$ \emph{(dominance)} iff $u$ is a prefix of $v$, i.e.\ the node
  $u$ is equal to or above the node $v$ in the tree.
  $u \wedgedom v$ iff $u \dom v$, and all symbols on the path from $u$
  to $v$ (not including $v$) are $\wedge$.
\end{definition}

@@

In the context of robust semantic processing, this is a desirable
feature, because it means that when we enrich an \rmrs\ obtained from
a shallow processor with more semantic information---such as the
relation symbols introduced by syntactic constructions such as
appositives, noun-noun compounds and free adjuncts---we don't change
the set of models; we only \emph{restrict} the set of models further
and further towards the semantic representation we are trying to
reconstruct.  Furthermore, it has been shown in the literature that a
dominance-constraint style semantics for underspecified
representations gives us more room to manoeuvre when developing
efficient solvers than an \mrs-style semantics
\cite{Althaus_etal:JoA}.

@@

\begin{definition} \label{defn:solved-form-of} The \emph{syntactic
    dominance relation} $D(\varphi)$ in an {\sc rmrs} $\varphi$ is the
  reflexive, transitive closure of the binary relation $$
\begin{array}{r c l}
\{(X,Y) & \mid &  \mbox{$\varphi$ contains $X \dom Y$ or}\\
&& \mbox{$ARG_S(X,Y)$ for some
    $S$}\}
\end{array}
$$

@@

Let's look at an example for all these definitions.  All the \rmrs s
presented in Section~\ref{sec:motivation} (replacing $\qeq$ by $\dom$)
are in solved form; this is least obvious for (\ref{ex:cat-erg}), but
becomes clear once we notice that no label is on the right-hand side
of two dominance atoms.  However, the model constructed in the proof
of Prop.~\ref{prop:solved-forms-are-satisfiable} looks a bit like
Fig.~\ref{fig:fat-black-cat}; both models are problematic in several
ways and in particular contain an unbound variable $y$ even though
they also contains a quantifier that binds $y$.  If we restrict the
class of models to those in which such variables are bound (as
\newcite{copestake:etal:2005} do), we can enforce that the quantifiers
outscope their bound variables without changing models of the \rmrs\
further---i.e., we add the atoms $h_3 \dom l_5$ and $h_8 \dom l_5$.
Fig.~\ref{fig:fat-black-cat} is no longer a model for the extended
\rmrs, which in turn is no longer in solved form because the label
$l_5$ is on the right-hand side of two dominance atoms.  Instead, it
has the following two solved forms:

@@

So far we have based the syntax and semantics of \rmrs\ on the
dominance relation from \newcite{egg:etal:2001} rather than the qeq
relation from \newcite{copestake:etal:2005}.  This is partly because
dominance is a weaker relation than qeq: If a dependency parser links
a determiner to a noun and this noun to a verb, then we can use
dominance but not qeq to represent that the predicate introduced by
the verb is outscoped by the quantifier introduced by the determiner
(see earlier discussion).  However, it is very straightforward to add
qeq to the syntax of the language and to extend the model theory
accordingly.  This extension adds a new atom $X \qeq Y$ to
Def.~\ref{defn:rmrs-syntax}, and $\tau,\alpha,g,\sigma$ will satisfy
$X \qeq Y$ iff $\alpha(X) \dom \alpha(Y)$, each node on the path is a
quantifier, and each step in the path goes to the rightmost child. All
the above propositions about solved forms still hold if ``dominance''
is replaced with ``qeq''.
